Many-core applications to online track reconstruction in HEP experiments

### 2013-10-22 papers/2013/CHEP/paper.tex
better conclusions, bib cleanups

  \begin{figure}[!tbp]
    \centering
    \includegraphics[width=0.9\linewidth]{figures/TimeCompZoom_MIC.png} 
    \caption{Algorithm-only comparison for timing as a function of the
      number of track fits: zoom in the low number of fits region. At
      low number of fits, the CPU performs better, due to start-up
      costs associated with data transfers.}
    \label{fig:algo_only_timing_zoom}
  \end{figure}

 \begin{figure}[!tbp]
   \centering
   \includegraphics[width=0.9\linewidth]{figures/Speedup_MIC}
   \caption{Speed-up with respect to a standard CPU (Intel Xeon
     E5630). The speed-ups plateau after about two million fits.
     %for
     %the GPUs (in red) while the Intel MIC still shows a small slope
     %beyond that point. 
     \fixme{Silvia: removed the comment about the different GPU/MI beahviour. These are the speedup values for the MIC: 0.07, 0.49, 1.79, 2.25, 2.32, 3.35, 4.38,4.74, 5.36, 5.16, 5.34 }}
   \label{fig:algo_only_speedup}
 \end{figure}

\subsubsection{Breakdown of computing time}
In Fig.~\ref{fig:breakdown} we show the fractional time spent in
various parts of the algorithm for the embarrassingly parallel algorithm 
(on Intel MIC) and the parallel algorithm (on NVIDIA Titan GPU), as a 
function of the number of fits. On both accelerator cards the fractional times
are constant over 500 input events, where computing resources are saturated. 
Unlike the MIC, the fit stage takes most of the time on the GPU: this 
could be caused by the intense memory access frequency intrinsic to
this part of the algorithm.  
\begin{figure*}[tbp]
\centering
\subfigure[]
{\label{fig:breakdown_MIC}
\includegraphics[width=0.45\linewidth]{figures/Mic5_perc_nk}}
\subfigure[]
{\label{fig:breakdown_GPU}
\includegraphics[width=0.45\linewidth]{figures/titan_perc}}
\caption{Breakdown of computing time for MIC (a) and the GTX Titan GPU
  (b). Red corresponds to unpacking, green corresponds to generating
  hit combinations, blue is the linearized track fit, and magenta
  corresponds to offloading (MIC only). For MIC, combinations and fitting
  take the same amount of time for large number of events. For GPU,
  fitting dominates. }
\label{fig:breakdown}
\end{figure*}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/datatransfer}
  \caption{Total latency (data transfer, copy to/from the GPU and data
    processing on the GPU) as a function of data buffer sizes, for
    three different levels of optimization of GPUDirect: v1.0,
    \textsc{cuda}-aware MPI and peer-to-peer. The smallest transferred
    data packed is 600 kB. \fixme{Below it says 600 kb is the place where
    P2P gets worse?}}
  \label{fig:xferlatency}
\end{figure}

In Fig.~\ref{fig:xferlatency} we show the total latency (data
transfer, copy to/from the GPU and data processing on the GPU) as a
function of data packet size when data are transferred using GPUDirect
v1.0, \textsc{cuda}-aware MPI and P2P. For the packet sizes over 600
kB \textsc{cuda}-aware MPI gives the best performance (see also~\cite{NSS2012} and~\cite{bib_mvapich})  \fixme{do we have this plot? Silvia: no, and will not have in time for the proceedings... but we can put a reference, see slides after page 47.}.

@@

\section{Conclusions}
We have examined a staged approach to using accelerator cards in a
hadron collider trigger application. We have demonstrated that in this
application, significant gains can be achieved with the
`embarrassingly parallel' approach on an Intel MIC
architecture, with the smallest amount of required changes to an
existing serial code base. However, better performance is achieved
with GPUs and a more complete event-level parallelization using
\textsc{cuda}.

>>>>

  \begin{figure}[!tbp]
    \centering
    \includegraphics[width=0.9\linewidth]{figures/TimeCompZoom_MIC.png} 
    \caption{Algorithm-only comparison for timing as a function of the
      number of track fits: zoom in the low number of fits region. At
      low number of fits, the CPU performs better, due to start-up
      costs associated with data transfers to the accelerator card.}
    \label{fig:algo_only_timing_zoom}
  \end{figure}

 \begin{figure}[!tbp]
   \centering
   \includegraphics[width=0.9\linewidth]{figures/Speedup_MIC}
   \caption{Speed-up with respect to a standard CPU (Intel Xeon
     E5630). The speed-ups plateau after about two million fits.}
   \label{fig:algo_only_speedup}
 \end{figure}

\subsubsection{Breakdown of computing time}
In Fig.~\ref{fig:breakdown} we show the fractional time spent in
various parts of the algorithm for the embarrassingly parallel algorithm 
(on Intel MIC) and the parallel algorithm (on NVIDIA Titan GPU), as a 
function of the number of fits. On both accelerator cards the fractional times
are constant over 500 input events, where computing resources are saturated. 
Unlike the MIC, the fit stage takes most of the time on the GPU: this 
could be caused by the intense memory access frequency intrinsic to
this part of the algorithm.  
\begin{figure*}[tbp]
\centering
\subfigure[]
{\label{fig:breakdown_MIC}
\includegraphics[width=0.45\linewidth]{figures/Mic5_perc_nk}}
\subfigure[]
{\label{fig:breakdown_GPU}
\includegraphics[width=0.45\linewidth]{figures/titan_perc}}
\caption{Breakdown of computing time for MIC (a) and the GTX Titan GPU
  (b). White corresponds to unpacking, green hash corresponds to
  generating hit combinations, solid blue is the linearized track fit,
  and magenta cross-hatch corresponds to offloading (MIC only). For
  MIC, combinations and fitting take the same amount of time for large
  number of events. For GPU, fitting dominates. }
\label{fig:breakdown}
\end{figure*}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/datatransfer}
  \caption{Total latency (data transfer, copy to/from the GPU and data
    processing on the GPU) as a function of data buffer sizes, for
    three different levels of optimization of GPUDirect: v1.0,
    \textsc{cuda}-aware MPI and P2P. The smallest transferred data
    packed is 600 kB. \fixme{Below it says 600 kb is the place where
      P2P gets worse?} \textsc{cuda}-aware MPI shows the best
    performance for larger packet size.}
  \label{fig:xferlatency}
\end{figure}

In Fig.~\ref{fig:xferlatency} we show the total latency (data
transfer, copy to/from the GPU and data processing on the GPU) as a
function of data packet size when data are transferred using GPUDirect
v1.0, \textsc{cuda}-aware MPI and P2P. For the packet sizes over 600
kB \textsc{cuda}-aware MPI gives the best performance (see
also~\cite{NSS2012} and~\cite{bib_mvapich})  \fixme{do we have this
  plot? Silvia: no, and will not have in time for the
  proceedings... but we can put a reference, see slides after page
  47. Peter: what page 47? I'm confused.}.

@@

\section{Conclusions}
We have implemented a full version of the CDF SVT tracking algorithm
on GPUs and Intel MIC. We examined a staged approach to using
accelerator cards in a hadron collider trigger application. We have
demonstrated that in this application, significant gains can be
achieved with the `embarrassingly parallel' approach on an Intel MIC
architecture, with the smallest amount of required changes to an
existing serial code base. However, better performance is achieved
with GPUs and a more complete event-level parallelization using
\textsc{cuda}.  We have updated latency studies and shown that for
larger packet sizes (greater than 600~kB), \textsc{cuda}-aware MPI
outperforms P2P, contrary to naive expectation.  Even at large packet
size, the data transfer takes an appreciable fraction of the total
algorithm time (about 20-25\%).
